---
title: Sustainability
subtitle: Trends in British Columbia's Population Size and Distribution
output:
  pdf_document:
    template: I:/SPD/Science Policy & Economics/State of Environment/_dev/templates/print_ver_template.tex
---

```{r set-options, echo=FALSE, cache=FALSE, warning=FALSE, message=FALSE}
library(sp) #for spatial files
library(maptools) #for fortify function
library(rgdal) #for spatial projection
library(bcmaps) #for BC boundary
library(dplyr) #for joining dataframes
library(raster) #for interesect and aggregate functions
library(envreportutils) #soe theme
library(ggplot2) #for plotting
library(RColorBrewer) #for colour palette
library(png) #for inserting image to plot
library(grid) #for creating grid graphic
library(knitr) #for outputing PDF

load("../tmp/popn_clean.RData")
opts_chunk$set(echo=FALSE, cache=FALSE, warning=FALSE, message=FALSE, error=TRUE,
               fig.height = 5.5)
read_chunk("../02_output.R")
```

```{r pre}
```
British Columbia's population growth is one of the fastest among provinces in Canada^1^. As the number of people living in the province increases so does pressure on the environment -- for example, through land use changes, increasing water demand, waste production, or emissions of pollutants. This indicator reports on the status and trends in British Columbia's population size and distribution for the province and among regional districts.

* **The population of British Columbia has grown continuously in the last century.** There were an estimated `r round(popn_bc[popn_bc$Year == 2015, "popn_million"], 1)` million people living in British Columbia as of 2015.

* **The majority of British Columbians live in the Greater Vancouver Regional District.** The 2015 estimated population size for Greater Vancouver is more than `r round(popn_sum[popn_sum$Regional.District == "Greater Vancouver", "Total"]/1000000, 1)` million people. The Capital Region District has the second largest population with approximately 400,000 people. The Fraser Valley and Central Okanagan have the third and fourth largest population in the province. 

* **Regional population sizes have increased in southern areas of B.C.** The Central Okanagan has seen the largest recent increase in population size -- the population more than doubled from 1986 to 2015. Other regions on the southwest coastal regions experienced various increases from 52% to 97%.

* **Regional population sizes have decreased in the northwest of B.C.** The Stikine has seen the largest recent decrease in population size. Population sizes in regional districts in central British Columbia, including Cariboo and Fraser-Fort George, experienced almost no change in estimated population size since 1986.
\newpage

## Long-term Change in British Columbia's Population Size (1867-2015)

```{r line_plot, fig.height = 4.5}
```

## B.C. Population Size and Density by Regional Districts (2015)

* British Columbia's population is not evenly distributed throughout the province.

* Greater Vancouver has the largest population size -- `r round(popn_sum[popn_sum$Regional.District == "Greater Vancouver", "Total"] / sum(popn_sum$Total) * 100, 0)`% of British Columbian's live in Greater Vancouver. The Capital and Fraser Valley Regional Districts have population sizes greater than 200,000 people, while the Stikine and Central Coast Regional Districts have the smallest regional population sizes in the province.

* Greater Vancouver has the highest population density with >800 people per square kilometer, followed by the Capital Regional District with 160 people per square kilometer. Most regional districts across the province have population densities of less than 10 people per square kilometer.

\newpage

```{r, fig.height = 7}
pal15 <- brewer.pal(5, "YlOrBr")

gv_barchart <- ggplot(data = popn_gv, aes(x = Regional.District, y = popn_thousand)) +
  geom_bar(stat = "identity", position = "identity", fill = pal15[5], 
           colour = "grey30", size = 0.2, alpha = 0.9) +
  labs(xlab("")) +
  labs(ylab("Population (*1000)")) +
  coord_flip() +
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradientn(colours = pal15) +
  theme_soe() +
  theme(axis.title.x = element_text(margin = margin(10, 0, 0, 0)),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 11),
        legend.position = "none",
        panel.grid = element_blank(),
        plot.margin = unit(c(0, 15, 15, 13.8), "mm"),
        text = element_text(family = "Verdana")) 

rest_barchart <- ggplot(data = popn_rest, aes(x = Regional.District, y = popn_thousand, fill = popn_thousand)) +
  geom_bar(stat = "identity", position = "identity", colour = "grey30", size = 0.3, alpha = 0.9) +
  coord_flip() +
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0), breaks = seq(0, 400, 80), limits = c(0, 400)) +
  scale_fill_gradientn(colours = brewer.pal(6, "YlOrBr")[1:2]) +
  theme_soe() +
  theme(axis.title = element_blank(),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 11),
        panel.grid = element_blank(),
        plot.margin = unit(c(10, 15, 5, 5), "mm"),
        legend.position = "none",
        text = element_text(family = "Verdana")) 
multiplot(rest_barchart, gv_barchart, cols = 1, heights = c(0.9, 0.2))
```

\newpage

```{r plot15, fig.height = 6}
```

\newpage

## B.C. Population Change by Regional District (1986-2015).
* British Columbia's population increased by `r round((popn_bc$Population[popn_bc$Year == 2015] - popn_bc$Population[popn_bc$Year == 1986])/popn_bc$Population[popn_bc$Year == 1986] * 100, 0)`% from 1986 to 2015, higher than Canada's 37% increase over the same period^1^.

* The Central Okanagan has seen the largest population increase of +`r popn_sum$Total_change[popn_sum$Regional.District == "Central Okanagan"]`% since 1986. Squamish-Lillooet, Fraser Valley, and Nanaimo Regional District have all experienced large increases in population size (> 80%).

* The Stikine has seen the largest decrease in population size (`r popn_sum$Total_change[popn_sum$Regional.District == "Stikine"]`%) since 1986. The Skeena-Queen Charlotte and Mount Waddington Regional Districts have also experienced recent declines in population size.

```{r change_map}
```

\newpage

## Methods
This indicator is based on data sourced from [BC Stats](http://bcstats.gov.bc.ca/Home.aspx), the central statistical agency of the Province of British Columbia. The [**R**](https://www.r-project.org/) code for repeating the analyses presented on this page is [available on GitHub](https://github.com/bcgov/bc_population_indicator).

## References and Other Useful Links

[BC Stats: Population Estimates](http://www.bcstats.gov.bc.ca/StatisticsBySubject/Demography/PopulationEstimates.aspx)

[Statistics Canada: Population and Demography](http://www5.statcan.gc.ca/subject-sujet/theme-theme.action?pid=3867&lang=eng&more=0)

^1^Annual Estimates of Population for Canada, Provinces and Territories, from July 1, 1971 to July 1, 2015. Newfoundland & Labrador Statistics Agency (published 09/29/15)(pdf):
<http://www.stats.gov.nl.ca/statistics/population/PDF/Annual_Pop_Prov.PDF>


## Data

\*By accessing these datasets, you agree to the license associated with each file, as indicated in parentheses below.

* British Columbia Annual Population Estimates (1867-2015):  <http://www.bcstats.gov.bc.ca/StatisticsBySubject/Demography/PopulationEstimates.aspx>
    - License: B.C. Crown Copyright: <http://www2.gov.bc.ca/gov/content/home/copyright>

* Total Population by Regional Districts in B.C. (1986-2015): <http://www.bcstats.gov.bc.ca/StatisticsBySubject/Demography/PopulationEstimates.aspx>
    - License: B.C. Crown Copyright: <http://www2.gov.bc.ca/gov/content/home/copyright>


Published and Available On-Line at Environmental Reporting BC (May
2016):
<http://www.env.gov.bc.ca/soe/indicators/sustainability/bc_population.html>

Email correspondence to: envreportbc@gov.bc.ca
